---
title: "Ecosystem Extent Map"
output: html_document
date: '2022-08-08'
---
Aim- to map changes in ecosystem extent by Moku in the main Hawaiian Islands (MHI) on an interactive map

This first notebook loads the raster data and carries out initial processing.
Exporting a cropped area of interest raster

Reference for processing raster Land Cover https://bookdown.org/mcwimberly/gdswr-book/raster-data-discrete-variables.html
---

```{r setup}
#library(leaflet)
library(raster)
library(rgdal)
library(tidyverse)
library(ggspatial)
```

```{r}
setwd("C:/Extent Map proj")
data_DIR <- "C:/all_data/remote sensing data"
raster_path <- file.path(data_DIR, "/LCMAP Land Cover data/lcmap_oahuWaianae_2015.tif")

#raster_path <- file.path(data_DIR, "/LCMAP Land Cover data/lcmap_2000_2011/LCMAP_HI_2004_V10_LCPRI.tiff")

#raster_path <- file.path(data_DIR, "/LCMAP Land Cover data/lcmap_2010_2020/LCMAP_HI_2020_V10_LCPRI.tiff")

# Oahu island only 
vector_path <- file.path(data_DIR, "/HIMARC/new data/Moku_NEW.shp")
```


First, import Land Cover/ Land-use raster, and vector layer to mask it with

```{r}
lu_raster <- raster(raster_path)
mask_vector <- readOGR(vector_path)
plot(mask_vector)
```
```{r}
sub_mask_vector <- mask_vector[mask_vector$NAME2 == "WAIANAE",]
plot(sub_mask_vector)
```


Mask LC raster with vector mask (ie. island/ Moku vector)
Note: projecting raster takes a long time, best to do it once and export it. Then just import it when you need it again.

```{r}

lu_raster_proj <- projectRaster(lu_raster, crs = "+proj=longlat +datum=WGS84", method = "ngb")
#writeRaster(lu_raster_proj,'lc_2004.tif',options=c('TFW=YES'), overwrite = TRUE)

# Or import previously projected raster

#lu_raster_proj <- raster("lc_2000.tif")
#lu_raster_proj_aft <- raster("lc_2020.tif")
sub_mask_vector <-spTransform(sub_mask_vector, CRS("+proj=longlat +datum=WGS84"))

#masked_raster <- mask(x = lu_raster_proj, mask = sub_mask_vector)
#plot(lu_raster_proj)


```

Crop the Raster Extent

```{r}
cropped <- crop(x = lu_raster_proj, y = extent(sub_mask_vector))
#cropped_b <- crop(x = lu_raster_proj, y = extent(sub_mask_vector))
#cropped_aft <- crop(x = lu_raster_proj_aft, y = extent(sub_mask_vector))
plot(cropped)
```


Import the "raster.df" function to convert cropped raster into df for ease of manipulation

```{r, eval=FALSE}
source("rasterdf.R")
```

This is the output for further analysis

```{r, eval=FALSE}
# Compare at the island level Land Cover change over two discrete years 

lc <- 8    # change this to landcover class of interest

cropped_df <- rasterdf(cropped_b)
cropped_df_aft <- rasterdf(cropped_aft)

# Number of pixels corresponding to developed land (ex.)
list(sum(cropped_df$value == lc), sum(cropped_df_aft$value == lc))

```

Loop to compile sum of pixels for individual Moku, and input into dataframe 
```{r}

oahu_moku <- list("EWA", "KONA OAH", "KOOLAUPOKO", "KOOLAULOA", "WAIALUA", "WAIANAE")
#kauai_moku <- list("MANA", "KONA KAU", "PUNA KAU", "KOOLAU KAU", "HALELEA", "NAPALI")

series_b <- c()

for (i in oahu_moku){
  cropped <- crop(x = lu_raster_proj, y = extent(sub_mask_vector[sub_mask_vector$NAME2 == i,]))
  cropped_df <- rasterdf(cropped)
  
  series_b <- c(series_b,sum(cropped_df$value == lc))    
}

series_aft <- c()

for (i in oahu_moku){
  cropped <- crop(x = lu_raster_proj_aft, y = extent(sub_mask_vector[sub_mask_vector$NAME2 == i,]))
  cropped_df <- rasterdf(cropped)
  
  series_aft <- c(series_aft,sum(cropped_df$value == lc))    
}

# Then with series of counts for each Moku, create a df

df <- data.frame(
  moku <- unlist(oahu_moku),
  lc <- series_b,
  lc_post <- series_aft)

colnames(df) <- c("moku", "lc")

print(df)

```




However, if you just want a .tif file this is the output.
Export cropped raster file

```{r}
#writeRaster(cropped,'lc_waianae_2015.tif',options=c('TFW=YES'))
```




# Produce nice LC map with Custom colourtable
This extension is for production of nice looking LC maps for reports, presentations etc. 
First import the .tif file from above output.

```{r}
cropped <- raster("lc_waianae_2015.tif")
cropped_df <- rasterdf(cropped)
```
  
```{r}
lc_codes <- c(1,2,3,4,5,6,8)
lc_names <- c("Developed", "Cropland", "Grass/Shrub", "Tree Cover", "Water", "Wetland", "Barren")
clrs <- c("#525252", "#AB6C28", "#CCB879", "#005a32", "#034e7b", "#4292c6", "#bdbdbd")

names(clrs) <- as.character(lc_codes)
clrs
```

```{r, eval=FALSE}
p <- ggplot(data = cropped_df) +
  geom_raster(aes(x = x, y = y, fill = as.character(value))) +
  ggtitle("WAIANAE, Oahu") +
  scale_fill_manual(name = "Land cover 2014",
                    values = clrs,
                    labels = lc_names,
                    na.translate = FALSE) +
  coord_sf(expand = F) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.background = element_rect(fill = "white"),
        plot.title = element_text(hjust = 0.5)
        )+
   theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.background = element_rect(fill = "white"),
        plot.title = element_text(hjust = 0.5)
        ) +
  annotation_north_arrow(height = unit(1, "cm"), 
                         width = unit(.65, "cm"),
                         pad_y = unit(1, "cm")) +
  annotation_scale(location = "bl")
p  

#ggsave(plot = p, filename = "oahu2000.png")
```


# Calculate area of each class






```{r}
library(sf)
```
```{r}
moku <- st_make_valid(st_read(vector_path))
moku <- st_transform(moku, CRS("+proj=longlat +datum=WGS84"))

mokuname <- "WAIANAE"
island <- "Oahu"

moku_crop <- moku[moku$NAME2 == mokuname,]
plot(st_geometry(moku_crop))
```


```{r}

moku_area <- sum(st_area(moku_crop[moku_crop$NAME2 == mokuname,]))

cropped_df <- cropped_df %>% filter(cropped_df$value > 0)
    
  agg_df<- cropped_df %>% 
    count(value) %>% 
    mutate('area m^2' = n * 100 * 100,
         'area km^2' = round((n * 100 * 100)/ 1e6, 2)) %>% 
    mutate('prop_cover' = `area m^2` / moku_area)
    
  agg_df$Land_cover <- agg_df$value  
  agg_df$Land_cover[agg_df$value == 1] <- "Developed"
  agg_df$Land_cover[agg_df$value == 2] <- "Cropland"
  agg_df$Land_cover[agg_df$value == 3] <- "Grass/Shrub"
  agg_df$Land_cover[agg_df$value == 4] <- "Tree Cover"
  agg_df$Land_cover[agg_df$value == 5] <- "Water"
  agg_df$Land_cover[agg_df$value == 6] <- "Wetland"
  agg_df$Land_cover[agg_df$value == 8] <- "Barren"
  agg_df$Island <- island
  
  agg_df
```


